### 2019-05-29 README.Rmd
initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# hotspots

<!-- badges: start -->
<!-- badges: end -->

The goal of hotspots is to ...

## Installation

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#library(hotspots)
## basic example code
```

## Code of Conduct

Please note that the 'hotspots' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.