### 2019-06-06 README.Rmd
added Makefile

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# hotspots

<!-- badges: start -->
[![Travis-CI Build Status](https://travis-ci.org/fboehm/hotspots.svg?branch=master)](https://travis-ci.org/fboehm/hotspots)

<!-- badges: end -->

The goal of hotspots is to ...

## Installation

Install from Dockerhub with this command:

```{bash, docker_pull, eval = FALSE}
docker pull fjboehm/hotspots
```

Alternatively, if you wish to install the package directly from Github, use `devtools`.

```{r, devtools_stuff, eval = FALSE}
install.packages("devtools")
devtools::install_github("fboehm/hotspots")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#library(hotspots)
## basic example code
```

## Build the manuscript `paper.Rmd`

To render the manuscript `paper.Rmd`, into html, tex, and pdf, use a shell to enter the following text from the 'hotspots' directory:

```{bash, docker_build, eval = FALSE}
docker build .
```

If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rmd`. Just type `make` in the shell:

```{bash, make, eval = FALSE}
make
```

## Code of Conduct

Please note that the 'hotspots' project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.