### 2019-06-24 analysis/paper/paper.Rnw
minor fixes to paper text

\documentclass[]{article}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{bbm}
\usepackage[left = 1in, right = 1in, top = 1.5in, bottom = 1in]{geometry}
%\usepackage[titletoc]{appendix}
\usepackage{amsmath}
\usepackage{hyperref}
\hypersetup{unicode=true,
%            pdftitle={Chapter 1},
%            pdfauthor={Frederick Boehm},
%            pdfborder={0 0 0},
            colorlinks=true,
            breaklinks=true}
\usepackage[yyyymmdd,hhmmss]{datetime}

\usepackage[
    backend=biber,
    style=chicago-authordate,
%    citestyle=authoryear,
    natbib=true, %citet and citep
    url=true,
    doi=true,
    eprint=true,
    maxbibnames=10,
    backref=true
]{biblatex}
\DeclareLanguageMapping{english}{english-apa}
\addbibresource{research.bib}
\usepackage{graphicx}
\usepackage{color, colortbl}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
%\usepackage{longtable}
%\usepackage[outdir=./figs/]{epstopdf}
%\usepackage{etoolbox} % for messing with page style & title pages
\usepackage{fancyhdr}
\setlength{\headheight}{15.2pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{}
\chead{}
\rhead{ \fancyplain{}{\thepage} } % put page number in upper right corner
\rfoot{Compiled on \today\ at \currenttime}
\renewcommand{\headrulewidth}{0pt} % remove top line
\renewcommand{\footrulewidth}{0pt} % remove bottom line
\renewcommand{\headrule}{}%

\title{Expression QTL hotspot dissection with mediation analysis and pleiotropy testing}
\author{Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman}
\date{\today}

\doublespacing

\begin{document}

\maketitle

<<setup, include=FALSE>>=
library(knitr)
opts_chunk$set(
concordance=TRUE,
echo = TRUE,
fig.lp = "fig:",
warning = FALSE,
message = FALSE)

options(tinytex.verbose = TRUE)
@

\listoftodos
\tableofcontents
\listoffigures
\listoftables
%\pagestyle{headings} % page numbers appear at bottom of page, centered!
%\markright{\hfill}
\section{Introduction}

Genetics studies in model organisms like mice can identify
genomic regions that affect complex traits, such as systolic blood pressure and body weight \citep{sax1923association,soller1976power,lander1989mapping,broman2009guide,jansen2007quantitative}. These genomic regions are called ``quantitative trait loci'' or ``QTL''.
A genome-wide QTL ``scan'' reveals
associations between genotypes and phenotypes by considering each position,
one at a time, as a candidate QTL for the trait of interest.
A region with strong evidence of association with a complex trait, then, defines a QTL (for that trait).
Because nearby markers have correlated
genotypes, a QTL in a two-parent cross often spans multiple megabases in length
and may contain more than a hundred genes.
Identification of the causal gene (for a given complex trait) from among those genes near the QTL is
challenging and may require costly and time-consuming experiments.
The growing need for greater QTL mapping resolution fueled development over the
last two decades of model organism multiparental populations for high-resolution QTL mapping \citep{de2017back,churchill2004collaborative,svenson2012high,huang2012multiparent,shivakumar2018soybean,huang2011analysis,kover2009multiparent,tisne2017identification,stanley2017genetic}.
With experimentalists now measuring tens of thousands of biomolecular traits
in multiparental populations, the systems genetics community needs multivariate
statistical tools to fully examine the large volumes of data \citep{keller2018genetic,chick2016defining}.
Identifying loci that affect multiple traits can aid in identifying biomolecular interactions and in clarifying complex trait genetic architecture. A QTL that affects more than one trait is called a ``pleiotropic'' QTL.
A test of pleiotropy vs. separate QTL is one multivariate statistical tool that will inform
complex trait genetics by enabling researchers to identify the number of
unique QTL in a genomic region of interest.

<<load-broman>>=
library(broman)
@

\section{Methods}

\citet{keller2018genetic} identified five expression trait hotspots in their
study of pancreatic islet gene expression in 378 Diversity Outbred mice. They
defined a hotspot to be a 4-Mb region that affects at least 100 expression
traits.

For each hotspot, we identified a set of local expression traits and a
set of nonlocal expression traits. All traits demonstrated LOD peaks that met
the genome-wide significance threshold, 7.18 \citep{keller2018genetic}.
The nonlocal expression traits arose from genes
on other chromosomes, while the local traits reflected expression levels for
genes on the same chromosome as the hotspot and within 4 Mb of the hotspot
center. We ignored transcripts for genes on the hotspot chromosome but not within 4 Mb of
the hotspot center.

\subsection{Pleiotropy testing}

For each hotspot, we paired each local trait with each nonlocal trait and performed a
two-dimensional, two-QTL scan over a genomic region that included the entirety of the
hotspot. We fitted bivariate linear mixed effects models at each ordered pair of markers in a two-dimensional grid \citep{boehm2019testing}.

\begin{equation}
vec(Y) = X vec(B) + vec(G) + vec(E)
\label{eq:mvlmm}
\end{equation}

From the two-dimensional scans, we calculated pleiotropy likelihood ratio test statistics for each
pair of traits \citep{jiang1995multiple,boehm2019testing, boehm2019thesis}. Because of the required
computing time for bootstrap analyses, we didn't obtain p-values for the pleiotropy tests. Instead,
we worked directly with the test statistic values.

\todo[inline]{copy equations from G3 paper}

\begin{equation}
LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
\label{eq:lrt}
\end{equation}

\begin{equation}
ll(\lambda_1, \lambda_2)
\end{equation}

\subsection{Mediation analysis}

For each hotspot and each local trait - nonlocal trait pair, we also performed a
mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait. Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) \citep{chick2016defining}.

\begin{equation}
Y = \alpha 1 + WC + E
\label{model1}
\end{equation}

\begin{equation}
Y = XB + WC + E
\label{model2}
\end{equation}

\begin{equation}
Y = \alpha 1 + WC + M\beta + E
\label{model3}
\end{equation}

\begin{equation}
Y = XB + WC + M\beta + E
\label{model4}
\end{equation}

In all mediation analyses, the nonlocal transcripts are the targets, while the local
transcripts serve as candidate mediators of the QTL-target relationship.
In Equations~\ref{model1} through \ref{model4}, $X$ is a matrix of founder allele dosages,
$B$ is a matrix
of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
effects, $M$ is a vector of values for a single candidate mediator, $\beta$ is the effect
of the mediator on the complex trait, 1 is a vector with all entries set to 1, $\alpha$
is the mean trait value, and $E$ is a vector of normally distributed random errors. \todo[inline]{better so say 'column matrix' or 'vector'? I hate to worry about this.}

\section{Results}

\section{Discussion}

\printbibliography

<<install_here>>=
install.packages("here", repos = "https://cloud.r-project.org")
@

\subsection{Colophon}

This report was generated on \Sexpr{Sys.time()} using the following computational environment and dependencies:

\begin{singlespacing}

<<colophon, cache = FALSE>>=
# which R packages and versions?
devtools::session_info()
@

The current Git commit details are:

<<git2r, cache = FALSE>>=
# what commit is this file at?
git2r::repository(here::here())
@

\end{singlespacing}

\end{document}