### 2019-07-15 analysis/paper/heatmaps-dendro-check.Rmd
added file to do checks on dendrograms & heatmaps

---
title: "Checking heatmaps and dendrograms"
author: "Frederick Boehm"
date: "7/14/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(hotspots)
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)

```

The questions that I want to address today deal with heatmaps and dendrograms. Basically, I want to ensure that I'm correctly sorting & ordering a data matrix before passing it to a heatmap function.

I need to check a few things. First, what is the ordering of rows of a matrix when I create a dendrogram? What is the ordering of rows when I create a heatmap (without attempting any manual reordering)?

Can I order the data matrix according to the dendrogram so that, when calling the heatmap function *while preventing the automated reordering of rows*, do I get the same ordering as in the dendrogram?

Other questions that I want to consider include the *method* for dendrogram creation. I also want to think about coloring schemes in heatmaps.

## Reading Chr 11 intermediate result files

Let's examine the Chr 11 files for pleiotropy results.

```{r}
pleio <- readRDS("_cache/Chr11_pleio.rds")
```