### 2019-07-16 analysis/paper/Chr11-heat.Rmd
added Rmd files for heatmaps

---
title: "Heatmaps of pleiotropy test statistics"
author: "Frederick Boehm"
date: "7/14/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(hotspots)
```

The questions that I want to address today deal with heatmaps and dendrograms. Basically, I want to ensure that I'm correctly sorting & ordering a data matrix before passing it to a heatmap function.

I need to check a few things. First, what is the ordering of rows of a matrix when I create a dendrogram? What is the ordering of rows when I create a heatmap (without attempting any manual reordering)?

Can I order the data matrix according to the dendrogram so that, when calling the heatmap function *while preventing the automated reordering of rows*, do I get the same ordering as in the dendrogram?

Other questions that I want to consider include the *method* for dendrogram creation. I also want to think about coloring schemes in heatmaps.

# Chr 11

```{r setup_continued}
hot_chr <- 11
hot_mid <- 71
```

```{r more_setup}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

## Reading Chr 11 intermediate result files

Let's examine the Chr 11 files for pleiotropy results.

```{r readRDS}
pleio_mixed_list <- readRDS("_cache/Chr11_pleio.rds")
pleio_local_list <- readRDS("_cache/Chr11_pleio_local.rds")
pleio_nonlocal_list <- readRDS("_cache/Chr11_pleio_nonlocal.rds")
med_list <- readRDS("_cache/Chr11_mediations.rds")
```

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r make_local_annot}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```

```{r make_nonlocal_annot}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```

```{r, make_local_annot2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```

```{r tabulate, message=FALSE, warning = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```

```{r make_pleio_tib}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r hier_clust}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
hc <- pm %>% 
  dist() %>%
  hclust()
hc %>%
  plot(cex = 0.2)
```

```{r}
hc$labels == rownames(pm)[hc$order]
```

```{r}
tibble(hc$labels, rownames(pm)[hc$order], hc$labels[hc$order], rownames(pm))
```

The tibble above shows the needed result. `hc$labels` is, by default definition, equal to `rownames(pm)`, while `rownames(pm)[hc$order]` gives the sequence of symbols that align with the leaves of the dendrogram.

```{r load_heatmaply}
library(heatmaply)
```

The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.

```{r heat-pleio}
heatmaply(pm, symm = TRUE)
```

```{r heat1}
heatmaply(percentize(pm), cexRow = 0.2, cexCol = 0.2, seriate = "OLO", symm = TRUE)
```

```{r heat2}
heatmaply(percentize(pm), cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE)
```

```{r heat3}
heatmaply(normalize(pm), 
          cexRow = 0.2, 
          cexCol = 0.2, 
          seriate = "OLO", 
          symm = TRUE, dendrogram = "both"
          )
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat4}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat5}
(pm_nonlocal > 3) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```

In the matrix of nonlocal - nonlocal pairs of pleiotropy test statistics, the fact that the maximum is only `r round(max(pm_nonlocal, na.rm = TRUE))` tells us that the heatmap of binary values is not going to be informative, since all or nearly all entries are less than any reasonable threshold.

Regarding possible transformations of the rows of the matrix, I don't really like transformations here. Transformations aim to preserve, in some sense, the distribution of values - test statistics, here - and I don't think that we should do that. I don't really care if a given row comes from a normal distribution, for example. I'm inclined to prefer heatmaps of the raw values.