### 2019-07-29 analysis/paper/intro.Rmd
fixed equation labels

# Introduction

Genetics studies in model organisms like mice can identify
genomic regions that affect complex quantitative traits, such as systolic blood pressure, body weight, and, more recently, biomolecular traits such as gene expression levels and protein concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining]. 
These genomic regions are called "quantitative trait loci" or "QTL".
A genome-wide QTL "scan" reveals
associations between genotypes and phenotypes by considering each position along the genome,
one at a time, as a candidate QTL for the trait of interest.
A region with strong evidence of association with a complex quantitative trait, then, defines a QTL (for that trait).
Because nearby markers have correlated
genotypes, a QTL in a two-parent cross often spans multiple megabases in length
and may contain more than a hundred genes.
Identification of a causal gene (for a given complex trait) from among those genes near the QTL is
challenging and may require costly and time-consuming experiments.
The growing need for greater QTL mapping resolution fueled development over the
last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].

Recent systems genetics studies have sought to identify QTL for thousands of biomolecular traits [@keller2018genetic;@chick2016defining].