### 2019-08-02 analysis/paper/results.Rmd
created discussion.Rmd

## Results

We studied five expression trait hotspots in pancreatic islets from 378 Diversity Outbred mice [@keller2018genetic]. Scatter plots of 
pleiotropy test statistic against mediation LOD difference for local - nonlocal trait pairs reveal 
some similarities for the five hotspots. All five scatter plots have few, if any, points in the upper 
right quadrant. This tells us that local-nonlocal trait pairs may have either high pleiotropy test 
statistics or high mediation LOD differences, but not both.

Examinating the ranges of pleiotropy test statistics within a hotspot also informs our conclusions. 
Some hotspots have maximum pleiotropy test statistics under 6 (Figure \@ref(fig:scatter-7)), while 
other hotspots have local-nonlocal pairs with pleiotropy test statistics over 15 (Figure 
\@ref(fig:scatter-2)).