### 2019-08-02 analysis/paper/discussion.Rmd
started Discussion section

## Discussion

We sought to answer the question of whether a single local trait is primarily responsible for each of 
five hotspots in pancreatic islets from Diversity Outbred mice. Our primary tools for probing this 
question are linear regression-based mediation analysis and pleiotropy testing. 
We