### 2019-08-02 analysis/paper/discussion.Rmd
moved discussion text from paper to discussion.Rmd

## Discussion

We sought to answer the question of whether a single local trait is primarily responsible for each of 
five hotspots in pancreatic islets from Diversity Outbred mice. Our primary tools for probing this 
question are linear regression-based mediation analysis and pleiotropy testing. 
We

Pleiotropy tests complement mediation analyses in the following manner. When considering a
collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
mediator when there is, in fact, a mediator among the candidates. However, when there is no
mediator, mediation analyses provide little information about the genetics of the complex trait.
Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
an answer about the number of underlying QTL. A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.

- lod difference threshold for declaration of "mediation". 
  - esp with regards to Chr 5 hotspot, where the keller mediator has a lot of traits with lod differences under 3, but above the 1.5 threshold.

- look at outputs of `lod_int()`, ie, the support intervals for all traits in a hotspot. Could plot interval against cluster order, for example.