### 2019-08-02 analysis/paper/intro.Rmd
updated intro

# Introduction

Genetics studies in model organisms like mice can identify
genomic regions that affect complex quantitative traits, such as systolic blood pressure, body 
weight, and, more recently, biomolecular traits such as gene expression levels and protein 
concentrations [@sax1923association;@soller1976power;@lander1989mapping;@broman2009guide;@jansen2007quantitative;@chick2016defining]. 
These genomic regions are called "quantitative trait loci" or "QTL".
A genome-wide QTL "scan" reveals
associations between genotypes and phenotypes by considering each position along the genome,
one at a time, as a candidate QTL for the trait of interest.
A region with strong evidence of association with a complex quantitative trait, then, defines a QTL (for that trait).
Because nearby markers have correlated
genotypes, a QTL in a two-parent cross often spans multiple megabases in length
and may contain more than a hundred genes.
Identification of a causal gene (for a given complex trait) from 
among those genes near the QTL is
challenging and may require costly and time-consuming experiments.
The growing need for greater QTL mapping resolution fueled development over the
last two decades of model organism multiparental populations for high-resolution QTL mapping [@de2017back;@churchill2004collaborative;@svenson2012high;@huang2012multiparent;@shivakumar2018soybean;@huang2011analysis;@kover2009multiparent;@tisne2017identification;@stanley2017genetic].

Recent systems genetics studies have sought to identify QTL for thousands of biomolecular traits [@keller2018genetic;@chick2016defining].

@boehm2019thesis argued for complementary roles for mediation analysis and pleiotropy testing in hotspot 
dissection. Here, we clarify this argument by presenting a study of five expression trait 
hotspots in a publicly available data set. We consider each hotspot one at a time. At each hotspot, we identify 
a set of local traits that have QTL within the hotspot. We also identify a set of nonlocal traits with QTL 
arising from the hotspot. We perform pleiotropy tests and mediation analyses for all local-nonlocal trait pairs. 
Graphical summaries of our results enable us to conclude that at least one of the five hotspots is due to multiple QTL, 
while the evidence at the four other hotspots is consistent with a single local trait affecting most or all of the 
nonlocal associated traits. The Chr 5 hotspot

## INTRO IDEAS

- why is it useful to study expression hotspots? 
- why are these two methods - pleiotropy tests and mediation analyses useful for studying hotspots? 
- who else has studied hotspots? And what did they do?