### 2019-08-03 analysis/paper/Rmd/abstract.Rmd
organized analysis/paper dir

# Abstract

We present our analysis of five expression trait hotspots in pancreatic islets from 378
Diversity Outbred mice.