### 2019-08-03 analysis/paper/preliminary/outline.Rmd

---
title: "Manuscript Outline"
author: "Frederick Boehm"
date: "7/22/2019"
output: html_document
---

## Figures list

1. pleiotropy v lod difference (one plot per hotspot)
1. pleiotropy v lod difference (multipanel figures: one panel per local trait per hotspot)
1. heatmaps of pleiotropy test statistics with mediation results (one plot per hotspot)

## Tables list

1. summary table with one row per hotspot, like Keller's supp table.
1. mediation counts per local trait (one table per hotspot). Each row is a local trait that mediates at least 4 nonlocal traits.