### 2019-10-02 analysis/paper/Rmd/methods.Rmd

# Methods

We used the publicly available data from @keller2018genetic to demonstrate our approach for 
expression trait hotspot dissection. @keller2018genetic measured genome-wide gene expression 
levels with RNA sequencing of pancreatic islets in 378 
Diversity Outbred mice. From the RNA sequencing counts, they obtained normalized gene 
expression levels for 21,771 transcripts, which they shared publicly via DataDryad [@dryad]. 
@keller2018genetic also included founder allele probabilities at 69,005 markers across the 
genome, covariates, and leave-one-chromosome-out estimated kinship matrices [@broman2019rqtl2; @yang2014advantages].

@keller2018genetic performed univariate QTL mapping for all expression traits. They
defined an expression trait hotspot to be a 4-Mb region that affects at least 100 nonlocal 
expression traits.
They identified five expression trait hotspots - on Chromosomes 2, 5, 7, 11, and 13 - in their data.

In analyzing the data from @keller2018genetic, we first established criteria for identifying local and 
nonlocal transcripts at each of the five hotspots. We defined a "local trait" to be a transcript whose corresponding gene lay within the hotspot. We focused on those local 
traits that strongly associated with QTL within the hotspot. To limit the scope of our analyses, 
we limited local traits to those with univariate LOD greater than 15.

We defined a "nonlocal trait" to be one with univariate evidence of association (LOD at least 7.18, the genome-wide threshold)
with a QTL in the hotspot [@keller2018genetic]. Additionally, we required that the nonlocal trait's gene not lie on the hotspot's chromosome. We ignored transcripts for genes on the hotspot chromosome 
but outside the hotspot interval.

For each hotspot, we performed both pairwise pleiotropy tests and linear regression-based 
mediation analyses. For every pair of traits (local or nonlocal) at a hotspot, we calculated a pleiotropy test 
statistic[@boehm2019testing]. Due to the computational burden, we didn't determine bootstrap p-values for these 
tests. For every local trait - nonlocal trait pair, we also performed a linear 
regression-based mediation analysis [@chick2016defining,@keller2018genetic]. We describe both procedures below.

## Pleiotropy testing

\begin{equation}
vec(Y) = X vec(B) + vec(G) + vec(E)
(\#eq:mvlmm)
\end{equation}

For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. For each test, we performed a
two-dimensional, two-QTL scan over a genomic region that included the entirety of the
hotspot. We fitted bivariate linear mixed effects models (Equations \@ref(eq:mvlmm), \@ref(eq:g), and \@ref(eq:e)) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].

\begin{equation}
G \sim MN_{n x 2}(0, K, V_g)
(\#eq:g)
\end{equation}

\begin{equation}
E \sim MN_{n x 2}(0, I_n, V_e)
(\#eq:e)
\end{equation}

and $G$ and $E$ are independent.

\begin{equation}
\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
(\#eq:Bhat)
\end{equation}

\begin{equation}
\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
(\#eq:Sigmahat)
\end{equation}

\begin{equation}
l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
(\#eq:loglik)
\end{equation}

For $n$ mice and two traits, $\Sigma$ is a $2n$ by $2n$ covariance matrix, while $V_g$ and $V_e$ are $2$ by $2$ covariance matrices (\@ref(eq:Sigmahat)). Estimates $\hat V_g$ and $\hat V_e$ are obtained via an expectation-maximization algorithm [@zhou2014efficient;@gemma2].

From the two-dimensional scan results, we calculated pleiotropy likelihood ratio test (LOD) statistics for each
pair of traits [@jiang1995multiple;@boehm2019testing;@boehm2019thesis]. Because of computing limitations, we didn't obtain bootstrap p-values for the pleiotropy tests. Instead,
we worked directly with the test statistic values.

\begin{equation}
LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
(\#eq:lod)
\end{equation}

## Mediation analysis

A mediation analysis in systems genetics begins with an identified QTL-expression trait 
association [@chick2016defining]. In our case, we choose a QTL that affects a nonlocal expression trait, then ask if any of 
the nearby local expression traits lie in the causal pathway from QTL to nonlocal trait.

A single mediation analysis involving one nonlocal trait and one local trait requires fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].

\begin{equation}
Y = \alpha 1 + WC + E
(\#eq:model1)
\end{equation}

\begin{equation}
Y = XB + WC + E
(\#eq:model2)
\end{equation}

\begin{equation}
Y = \alpha 1 + WC + M\beta + E
(\#eq:model3)
\end{equation}

\begin{equation}
Y = XB + WC + M\beta + E
(\#eq:model4)
\end{equation}

In all of our mediation analyses, the nonlocal traits are the targets (Y), while the local
transcripts serve as candidate mediators (M) of the QTL-target relationship. 
We only considered as potential mediators those local traits with strong univariate LODs.
In Equations~\ref{eq:model1} through \ref{eq:model4}, $X$ is a matrix of founder allele dosages,
$B$ is a matrix
of founder allele effects, $W$ is a matrix of covariates, $C$ is a matrix of covariate
effects, $M$ is a vector of values for a single candidate mediator (a local expression trait, in our case), $\beta$ is the effect
of the mediator on the complex trait (after adjusting for other covariates in the model), $1$ is a vector with all entries set to 1, $\alpha$
is a parameter that corresponds to the mean trait value (after adjusting for other covariates in the model), and $E$ is a vector of normally distributed random errors.

We included four covariates in our models: three wave indicators and sex. The wave indicators are binary vectors to account for wave membership. @keller2018genetic phenotyped the mice in four waves of nearly 100 mice per wave.

## Visualizations

We created scatter plots with the `ggplot2` package [@ggplot2] for the R
statistical computing environment [@r]. We plotted pleiotropy test statistics against 
mediation LOD differences for all local-nonlocal pairs at each hotspot. We highlighted in blue 
those points that correspond to pairs involving the putative mediator that @keller2018genetic 
identified. We used the same results in plots of pleiotropy test statistics against mediation LOD difference for analyses involving a single local trait.

We created heatmaps to examine patterns among pleiotropy test results in each hotspot. Our analysis used both 
local and nonlocal traits.
We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
trait and each column was an expression trait. 
Excepting the cells on the diagonal, each cell, then, contained the pleiotropy test statistic
from the analysis involving the row's expression trait and the column's expression trait.
To create a heatmap, we first performed two hierarchical clusterings of rows. 
One hierarchical clustering involved only the rows for local traits, while the other contained only 
rows corresponding to nonlocal traits.
We calculated the Euclidean distances between all row pairs (in the respective 
submatrices) before clustering. Hierarchical clustering was performed using the iterative complete 
linkage method. Initially, each row is assigned to its own cluster. The iterative 
algorithm then combines the two most similar clusters at each step. Distances between 
clusters were computed with the Lance-Williams dissimilarity update formula for complete 
linkage [@lance1967general]. Lastly, we arranged the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
We then combined the
local and nonlocal traits into a single heatmap for each hotspot. 
To achieve coloring contrasts in the heatmaps, we
set the color spectrum limits at 0 and 5. 
We annotated rows with six pieces of information: 1. mediation LOD differences, 2. mediation 
LOD difference proportions, 3. univariate QTL peak positions, 4. estimated allele effects 
vector at the univariate peak position, 5. univariate LOD score, and 6. whether the row 
corresponds to a local trait. For clarity of presentation, we also indicated those columns that  
correspond to a local traits. We didn't examine the possibility that a local trait might 
mediate the relationship between a QTL and another local trait.