### 2020-03-25 analysis/paper/Rmd/hotspot-one-chromosome-TIMBR-Chr2.Rmd
set up for using child processes with TIMBR analyses

---
title: "QTL mapping and TIMBR for Chr 2"
author: "Frederick J. Boehm"
date: "03/18/2020"
output: html_document
params:
  lastmod: !r lubridate::now()
---

Last modified: `r params$lastmod`.

```{r setupx, include=FALSE, dev="CairoPNG"}
.libPaths("~/work/libs")
knitr::opts_chunk$set(dev="CairoPNG")
```

```{r setup}
library(qtl2hotspots)
ult <- 15 # univariate lod threshold
```

```{r load_tidyverse}
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```

```{r saveRDS}
if (!file.exists(here::here("analysis", "data", "derived_data", "genoprobs.rds"))){
  download.file("https://datadryad.org/stash/downloads/file_stream/63306",
              destfile = here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
  load(here::here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
  saveRDS(dataset.islet.rnaseq$lod.peaks, here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
  saveRDS(dataset.islet.rnaseq$annots, here::here("analysis", "data", "derived_data", "annots.rds"))
  saveRDS(dataset.islet.rnaseq$expr, here::here("analysis", "data", "derived_data", "expr.rds"))
  saveRDS(K, here::here("analysis", "data", "derived_data", "kinship.rds"))
  saveRDS(map, here::here("analysis", "data", "derived_data", "map.rds"))
  saveRDS(genoprobs, here::here("analysis", "data", "derived_data", "genoprobs.rds"))
  saveRDS(dataset.islet.rnaseq$covar, here::here("analysis", "data", "derived_data", "covar.rds"))
}
```

```{r readRDS}
lod_peaks <- readRDS(here::here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here::here("analysis", "data", "derived_data", "annots.rds"))
expr <- readRDS(here::here("analysis", "data", "derived_data", "expr.rds"))
probs <- readRDS(here::here("analysis", "data", "derived_data", "genoprobs.rds"))
K <- readRDS(here::here("analysis", "data", "derived_data", "kinship.rds"))
covar <- readRDS(here::here("analysis", "data", "derived_data", "covar.rds"))
map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
```

```{r}
hot_chr <- 2
hot_mid <- 165.5
keller_mediator <- "Hnf4a"
```

```{r}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
```

```{r, define_local_annot}
local_annot <- lod_peaks %>%
  dplyr::filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  dplyr::filter(!duplicated(annot.id)) %>%
  dplyr::filter(lod >= ult) %>%
  dplyr::mutate(local_tx = chr == hot_chr) %>%
  dplyr::filter(local_tx)
```

```{r, define_nonlocal_annot}
nonlocal_annot <- lod_peaks %>%
  dplyr::filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  dplyr::filter(!duplicated(annot.id)) %>%
  dplyr::filter(lod >= 7.18) %>%
  dplyr::mutate(local_tx = chr == hot_chr) %>%
  dplyr::filter(!local_tx) %>%
  dplyr::filter(!is.na(hotspot))
```

```{r, local_annot2}
local_annot2 <- local_annot %>%
  dplyr::filter(middle <= hot_upper, middle >= hot_lower) %>% 
  dplyr::arrange(desc(lod))
```

```{r hot_expr}
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
```

```{r, hot_annot-hot_chr}
hot_annot <- local_annot2 %>%
  dplyr::bind_rows(nonlocal_annot) %>%
  dplyr::select(annot.id, symbol)
```

```{r, scan1}
fn <- here::here("analysis", "data", "derived_data", "Chr2_scan1_out.rds")
if (file.exists(fn)) {
  s1out <- readRDS(fn)
} else {
  s1out <- qtl2::scan1(genoprobs = probs, 
                       pheno = cbind(hot_local, hot_nonlocal), 
                       kinship = K[["{{hot_chr}}"]], 
                       addcovar = covar, 
                       reml = TRUE, 
                       cores = 1
                       )
  saveRDS(s1out, fn)
}
```

```{r allele-effects}
fn <- here::here("analysis", "data", "derived_data", "Chr2_scan1coef_out.rds")
if (file.exists(fn)) {
  s1coef_out <- readRDS(fn)
} else {
s1coef_out <- purrr::map(.x = as_tibble(cbind(hot_local, hot_nonlocal)), 
           .f = function(pheno) {
             ph <- as.matrix(pheno)
             rownames(ph) <- rownames(expr)
             qtl2::scan1coef(genoprobs = list(probs[[hot_chr]]), 
                             pheno = ph, 
                             kinship = K[[hot_chr]], 
                             addcovar = covar, 
                             reml = TRUE, 
                             cores = 1
                             )
           }
                )
saveRDS(s1coef_out, fn)
}
```

```{r, get-allele-effects}
# add index values for peak position
local_nonlocal_annot <- dplyr::bind_rows(local_annot2, nonlocal_annot) %>%
  mutate(peak_index = match(table = names(map[[hot_chr]]), x = marker.id)) 
```

### TIMBR analyses

```{r, timbr-setup}
tr <- cbind(hot_local, hot_nonlocal)
prior_M <- list(model.type = "crp",
                prior.alpha.type = "gamma",
                prior.alpha.shape = 2.3009322,
                prior.alpha.rate = 0.7488104
                )
library(TIMBR)
data(mcv.data) # get A matrix
peaks <- qtl2::find_peaks(s1out, map)
```

```{r}
fn <- "attie_DO500_genoprobs_v5_36state_chr2.rds"
gp <- readRDS(here::here("analysis", "data", "derived_data", fn))
```

```{r, timbr}
pheno <- cbind(hot_local, hot_nonlocal)
set.seed(3411192) # to ensure getting the same samples with TIMBR
t_out <- peaks %>%
  dplyr::filter(chr == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::mutate(marker_index = purrr::map2_int(.x = pos, .y = chr, .f = function(x, y) which(map[[y]] == x))) %>%
  dplyr::mutate(t_out = purrr::pmap(.l = list(lodcolumn, chr, marker_index), .f = function(lodcolumn, chr, marker_index){
    phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]
    phe <- phe_pre[!is.na(phe_pre), , drop = FALSE]
    pr_pre <- gp[[1]][, , marker_index] # use first (and only!) entry in gp
    pr <- pr_pre[rownames(pr_pre) %in% rownames(phe), ]
    cov_pre <- cbind(1, covar)
    cov_pre2 <- cov_pre[rownames(cov_pre) %in% rownames(phe), ]
    cov <- qtl2:::drop_depcols(cov_pre2)
    TIMBR::TIMBR(y = phe, 
                 prior.D = list(P = pr, 
                                A = mcv.data$prior.D$A, # works for autosomes
                                fixed.diplo = FALSE), 
                 prior.M = prior_M, 
                 samples = 10000, 
                 samples.ml = 10000,
                 Z = cov, 
                 verbose = TRUE
                 )  
}))
saveRDS(t_out, here::here("analysis", "data", "derived_data", "timbr-chr2-10ksamples.rds"))
```

```{r}
t_out <- readRDS(here::here("analysis", "data", "derived_data", "timbr-chr2-10ksamples.rds")) %>%
  dplyr::rename(timbr = t_out)
```