### 2020-03-26 analysis/paper/Rmd/hotspot-one-chromosome-timbr.Rmd
added dplyr:: prefix; deleted unneeded code

## Chr {{hot_chr}}

```{r, def-{{hot_chr}}}
hot_lower <- {{hot_mid}} - 2
hot_upper <- {{hot_mid}} + 2
```

```{r, define_local_annot-{{hot_chr}}}
local_annot <- lod_peaks %>%
  dplyr::filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  dplyr::filter(!duplicated(annot.id)) %>%
  dplyr::filter(lod >= ult) %>%
  dplyr::mutate(local_tx = chr == {{hot_chr}}) %>%
  dplyr::filter(local_tx)
```

```{r, define_nonlocal_annot-{{hot_chr}}}
nonlocal_annot <- lod_peaks %>%
  dplyr::filter(chrom == {{hot_chr}}, pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  dplyr::filter(!duplicated(annot.id)) %>%
  dplyr::filter(lod >= 7.18) %>%
  dplyr::mutate(local_tx = chr == {{hot_chr}}) %>%
  dplyr::filter(!local_tx) %>%
  dplyr::filter(!is.na(hotspot))
```

```{r, local_annot2-{{hot_chr}}}
local_annot2 <- local_annot %>%
  dplyr::filter(middle <= hot_upper, middle >= hot_lower) %>% 
  dplyr::arrange(desc(lod))
```

```{r hot_expr-{{hot_chr}}}
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
```

```{r, hot_annot-{{hot_chr}}}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) 
```

### TIMBR analyses

```{r, timbr-setup}
prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
  prior.alpha.type = "gamma",
  prior.alpha.shape = 1,
  prior.alpha.rate = 2.333415)
prior_M2 <- list(model.type = "crp",
                prior.alpha.type = "gamma",
                prior.alpha.shape = 2.3009322,
                prior.alpha.rate = 0.7488104
                )
library(TIMBR)
data(mcv.data) # get A matrix
```

```{r}
fn <- "attie_DO500_genoprobs_v5_36state_chr{{hot_chr}}.rds"
gp <- readRDS(here::here("analysis", "data", "derived_data", fn))
```

```{r}
library(future)
options(future.globals.maxSize= 10 * 1024 ^ 3)
```

```{r}
pheno <- cbind(hot_local, hot_nonlocal)
set.seed(3411192) # to ensure getting the same samples with TIMBR
future::plan("multiprocess")
t_out <- hot_annot %>%
  dplyr::filter(pos <= hot_upper, pos >= hot_lower) %>%
  dplyr::mutate(marker_index = purrr::map_int(.x = pos, .f = function(x) which(map[[{{hot_chr}}]] == x))) %>%
  dplyr::mutate(t_out = furrr::future_pmap(.l = list(lodcolumn, marker_index), .f = function(lodcolumn, marker_index){
    phe_pre <- pheno[ , colnames(pheno) == lodcolumn, drop = FALSE]
    phe <- phe_pre[!is.na(phe_pre), , drop = FALSE]
    pr_pre <- gp[[1]][, , marker_index] # use first (and only!) entry in gp
    pr <- pr_pre[rownames(pr_pre) %in% rownames(phe), ]
    cov_pre <- cbind(1, covar)
    cov_pre2 <- cov_pre[rownames(cov_pre) %in% rownames(phe), ]
    cov <- qtl2:::drop_depcols(cov_pre2)
    TIMBR::TIMBR(y = phe, 
                 prior.D = list(P = pr, 
                                A = mcv.data$prior.D$A, # works for autosomes
                                fixed.diplo = FALSE), 
                 prior.M = prior_M1, 
                 samples = 10000, 
                 samples.ml = 10000,
                 Z = cov, 
                 verbose = TRUE
                 )  
}))
dim(t_out)
saveRDS(t_out, here::here("analysis", "data", "derived_data", "timbr-chr{{hot_chr}}-10ksamples-M1.rds"))
```